---
title: "Good & Bad Figure Assignment"
author: "Robert Dellinger"
date: "3/13/2022"
output: html_document
---

#Load Library
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

require(ncdump)
require(akima)
require(ncdf4)
require(dplyr)
require(rerddap)
require(sf)
require(tidyverse)
require(lubridate)
require(oce)
require(ocedata)
require(RColorBrewer)
require(wior)
require(rgeos)
require(ggsn)
require(ggplot2)
require(ggrepel)
require(ggmap)
require(marmap)
require("devtools")
```


```{r data, include=FALSE}
#Load Data
data("coastlineWorld") # loading in coastline shapes
data("coastlineWorldMedium")
data("coastlineWorldFine")

```


#Bad Plot

The bad plot has no labels for the axis and the axis has numbers overlapping one another. This plot also is distorted and the grid lines are messed up showing a projection that is inaccurate. The points on the map are far too large and the writing is eligible. Further, the color scheme for the map does not accurately depict the color used for ocean bathymetry. 


```{r Setting Up the Bad Plot, warning=FALSE}

par(mar=c(2, 2, 1, 1)) #setting margins
lonlim = c(-125,-110) #setting lat and long lim
latlim = c(32,40)

mpbad <- function() {
mapPlot(coastlineWorld, 
        projection="+proj=stere",#setting projection
        col="purple", 
        longitudelim=lonlim, 
        latitudelim=latlim, 
        lonlabels=TRUE,
        latlabels=TRUE,
        grid = c(.25, .25), # grid size
        geographical=2)
  
}

bathy <- function() {  #coordinates for bathymetry and setting function 
b <- as.topo(getNOAA.bathy(-180, 0, 30, 50, keep=TRUE))
mpbad() 
mapImage(b, col=oceColorsViridis, breaks=seq(-4000, 0, 100)) #depth levels 
mapImage(b, col=oceColorsViridis, breaks=seq(-4000, 0, 100), filledContour = TRUE) #color contour
mapPolygon(coastlineWorldFine, col="purple")
}

```


```{r, warning=FALSE}

mpbad()
bathy()

## create points name with their respective locations
  
stations1 <- data.frame(name = c("BML", "HMS", "MSB", "CDMSP", "LCSB","SC"), lat = c(38.33, 36.62, 34.36, 33.59, 34.06, 33.55), lon = c(-123.05, -121.90, -119.44, -117.87, -118.93, -117.80))

## define the margin of the plot
par(mar=c(2, 2, 3, 3))

## make a base and bathymetry
bathy()

mapGrid(.1,.1)

## ## overlay the point layer
mapPoints(longitude = stations1$lon,
          latitude = stations1$lat, 
          pch = 20, 
          cex = 5.00)

## add station name on the map
mapText(longitude = stations1$lon+2,
        latitude = stations1$lat+2,
        labels = stations1$name, font=5)

```

#Good Plot
ggggggggggggggvhjgvbc vbghcv fgbv lo 
The good plot has proper access labels clearly distinguishing latitude and longitude with propely sized anf formated gridelines. The points on the graph are clear and they are also in a legible font. The bathymetry of the ocean floor and the map are properly colored for bathymetry and projected properly onto the map. 


```{r, Setting Up the Good Plot, warning=FALSE}

par(mar=c(2, 2, 1, 1))
lonlim = c(-125,-110)
latlim = c(32,40)

mp <- function() {
mapPlot(coastlineWorldFine,  #finescale data
        projection="+proj=mill", # projection type
        col="darkolivegreen3", 
        longitudelim=lonlim, 
        latitudelim=latlim, 
        lonlabels=TRUE,
        latlabels=TRUE,
        grid = c(1, 1), # gridsize
        geographical=4)
}

bathy <- function() {
b <- as.topo(getNOAA.bathy(-180, 0, 30, 50, keep=TRUE)) #coords for bathy
mp()
mapImage(b, col=oceColorsGebco, breaks=seq(-4000, 0, 500))
mapImage(b, col=oceColorsGebco, breaks=seq(-4000, 0, 500), filledContour = TRUE) #oce gradient and filled contour for bathymetry 
mapPolygon(coastlineWorldFine, col="darkolivegreen3") #green topography
}
```


```{r, warning=FALSE}

mp()
bathy()

## create points name with their respective locations
  
stations1 <- data.frame(name = c("BML", "HMS", "MSB", "CDMSP"), lat = c(38.33, 36.62, 34.36, 33.59), lon = c(-123.05, -121.90, -119.44, -117.87))

stations2 <- data.frame(name = c("LCSB","SC"), lat = c(34.06, 33.55), lon = c(-118.93, -117.80))

## define the margin of the plot
par(mar=c(2, 2, 1, 1))

## make a base and bathymetry
bathy()

mapGrid(1,1)

## ## overlay the point layer
mapPoints(longitude = stations1$lon,
          latitude = stations1$lat, 
          pch = 20, 
          cex = 1.00)

mapPoints(longitude = stations2$lon,
          latitude = stations2$lat, 
          pch = 20, 
          cex = 1.00)

## add station name on the map
mapText(longitude = stations1$lon+0.65, 
        latitude = stations1$lat+0.5,
        labels = stations1$name, font=2)

mapText(longitude = stations2$lon-0.30,
        latitude = stations2$lat-0.45,
        labels = stations2$name, font=2)

```



