---
title: "Week 8: Functions"
author: "Robert Dellinger"
date: "3/8/2022"
output: html_document
---



```{r, Load Library, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(PNWColors)

```


```{r, Read in Data}

df <- tibble::tibble(
  a = rnorm(10), # draws 10 random values from a normal distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
head(df)


```
"There are three key steps to creating a new function:

You need to pick a name for the function. Here I’ve used rescale01 because this function rescales a vector to lie between 0 and 1.

You list the inputs, or arguments, to the function inside function. Here we have just one argument. If we had more the call would look like function(x, y, z).

You place the code you have developed in body of the function, a { block that immediately follows function(...).

return tells us what values we want returned when you run the function" 
(Silbiger, Lecture 8,  2022)

```{r, Making a Function}


fahrenheit_to_celsius <- function() {
}

fahrenheit_to_celsius <- function() { 
  temp_C <- (temp_F - 32) * 5 / 9
}

fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9 
}

fahrenheit_to_celsius <- function(temp_F) { 
  temp_C <- (temp_F - 32) * 5 / 9 
  return(temp_C)
}

fahrenheit_to_celsius(32)


```


```{r,Think, Pair, Share}

c_to_k <- function(temp_qC){
temp_K <- (temp_C +273.15)
return(temp_K)
}

c_to_k(0)


```

```{r, Making Plots into a Function, warning=FALSE, message=FALSE}

myplot<-function(data, x, y){ 
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

#layering the plot
myplot(x = body_mass_g, 
       y = flipper_length_mm)+
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")


```

```{r, Simple If/Else Functions}

a <- 4
b <- 5

if (a > b) { # my question
  f <- 20 # if it is true give me answer 1
    } else { # else give me answer 2
  f <- 10
}

```

```{r, If/Else Function}

myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
if(lines==TRUE){
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
else{
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
} 


mplot
```

