---
title: "Week 12: Intro to Models"
author: "Robert Dellinger"
date: "4/28/2022"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: cerulean
---

## Load Library

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../output/')

library(tidyverse)
library(here)
library(performance) 
library(modelsummary)
library(tidymodels)
#library(purr)

```

#Lab
timeseries of temperature and light data collected in tide pools in Oregon by Jenn Fields. 
- bring in all 4 files and calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool. 
- Use both a for loop and map() functions in your script. (Basically, do it twice). Due Tuesday at 1pm. 

# Download Data 
```{r}


TP1 <- read_csv(here("Week_13","Data", "TP1.csv"))
TP2 <- read_csv(here("Week_13", "Data", "TP2.csv"))
TP3 <- read_csv(here("Week_13", "Data", "TP3.csv"))
TP4 <- read_csv(here("Week_13", "Data", "TP4.csv"))

data.full <- full_join(TP1, TP2)
data2.full <- full_join(TP3, TP4)
dataU <- full_join(data.full, data2.full)

#glimpse(dataU)
```

# Creating a loop 

```{r Loop}

data.file <- here("Week_13", "data") # Creating a path 
files <- dir(path = data.file, pattern = ".csv") # Direct path 
filesU <- dir(path = data.file, pattern = ".csv", full.names = TRUE) # full name for map function

# empty data frame 
temp.light.dataframe <- data.frame(matrix(nrow = length(files), ncol = 5)) 
colnames(temp.light.dataframe) <- c("file", "mean_temp", "sd_temp", "mean_light", "sd_light") #naming columns in the data set 

for (i in 1:length(files)) { 
  Rdata <- read_csv(here(data.file, files[i])) #reading in files
  temp.light.dataframe$file[i]<-files[i] # new dataframe
  temp.light.dataframe$mean_temp[i]<- mean(Rdata$Temp.C, na.rm = TRUE) #calculate temp mean
  temp.light.dataframe$sd_temp[i]<- sd(Rdata$Temp.C, na.rm = TRUE) # calculate sd 
  temp.light.dataframe$mean_light[i]<- mean(Rdata$Intensity.lux, na.rm = TRUE) # light mean
  temp.light.dataframe$sd_light[i]<- sd(Rdata$Intensity.lux, na.rm = TRUE) # calculate sd
}

glimpse(temp.light.dataframe)
```
# Creating a map

```{r Map Using Purr}

# using purrr package 
#map <- filesU %>% # new data frame
 # set_names() %>% 
  #map_df(read_csv, .id = "file") %>% 
#  mutate(file=gsub(".csv","",basename(file))) %>% # removes name/path
 # group_by(file) %>% # groups for next function
  #summarise(mean_temp = mean(Temp.C, na.rm = TRUE), # calculate mean
    #        std_temp= sd(Temp.C, na.rm = TRUE), # calculate standard deviation
     #       mean_light = mean(Intensity.lux, na.rm = TRUE), #calculate mean
      #      std_light= sd(Intensity.lux, na.rm = TRUE)) # calculate standard deviation)


```

