---
title: "Week_5_Lab2_Plot"
author: "Robert Dellinger"
date: "2/24/2022"
output: html_document
---

```{r Library, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)


```


```{r Read in Conductivity and Depth Data}

conductivity.data <- read_csv(here("Week_5","data", "CondData.csv")) # read in conductivity data
  
view(conductivity.data)

depth.data <- read_csv(here("Week_5","data", "DepthData.csv"))
# read in depth data 
  
view(depth.data)
```

```{r Convert Date Columns Appropriately}

conductivity.data <- conductivity.data %>% 
  mutate(datetime=mdy_hms(depth)) %>% #mutate to change a column
  rename("Date" = datetime) %>% 
  drop_na() %>%  # drop na's
  select(Date, TempInSitu, Serial, SalinityInSitu_1pCal)


?rename


depth.data %>% 
  mutate(datetime=mdy_hms(date)) %>% #mutate to change a column
  drop_na () # drop na's
```

```{r Round the Conductivity Data to the Nearest 10 Seconds}

```


```{r Join the Dataframes Together}
Join the two dataframes together (in a way where there will be no NAs... i.e. join in a way where only exact matches between the two dataframes are kept)

```



take averages of date, depth, temperature, and salinity by minute (hint: easiest way to do this is to make a new column where the hours and minutes are extracted)
Make any plot using the averaged data
Do the entire thing using mostly pipes (i.e. you should not have a bunch of different dataframes). Keep it clean.
Don't forget to comment!
Save the output, data, and scripts appropriately


today() #the current date

today(tzone = "GMT") #date by time zone

am(now()) # is it morning?

leap_year(now()) # is it a leap year?

#Dates must be a character (y = year, m = month, d = day)

today() #the current date

today(tzone = "GMT") #date by time zone

am(now()) # is it morning?

leap_year(now()) # is it a leap year?

#Examples of the same date in different formats

ymd("2021-02-24")
mdy("02/24/2021")
mdy("February 24 2021")
dmy("24/02/2021")

```


```{r Date & Time Specifications w/ Lubridate}

ymd_hms("2021-02-24 10:22:20 PM")

mdy_hms("02/24/2021 22:22:20")

mdy_hm("February 24 2021 10:22 PM")

# make a character string
datetimes<-c("02/24/2021 22:22:20",
             "02/25/2021 11:21:10",
             "02/26/2021 8:01:52")

# convert to date
datetimes <- mdy_hms(datetimes)

month(datetimes)

month(datetimes, label = TRUE)

month(datetimes, label = TRUE, abbr = FALSE) #Spell it out

day(datetimes) # extract day

wday(datetimes, label = TRUE) # extract day of week

#convert to time

hour(datetimes)
minute(datetimes)
second(datetimes)

#adding time and days
datetimes + hours(4) # this adds 4 hours

datetimes + days(2) # this adds 2 days

#Rounding dates and time
round_date(datetimes, "minute") # round to nearest minute

round_date(datetimes, "5 mins") # round to nearest 5 minute


```

```{r Think Pair Share}

conductivity.data <- read_csv(here("Week_5","data", "CondData.csv"))%>% 
mutate(datetime=mdy_hms(depth)) %>% #use mutate to change a column
  drop_na () # drop na's

view(conductivity.data) 

```

```{r R package: Catter Plots}

library(devtools)
install_github("Gibbsdavidl/CatterPlots") # install the data

library(CatterPlots)
x <-c(1:10)# make up some data
y<-c(1:10)
catplot(xs=x, ys=y, cat=3, catcolor='blue')

```


```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
